
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r }
library(dplyr)
library(sjlabelled)
library(readr)
titanic_test <- read_csv("/Users/cordeliaradin/Desktop/datasets/titanic_test.csv")
titanic_train <- read_csv("/Users/cordeliaradin/Desktop/datasets/titanic_train.csv")
df1<-titanic_train
df2<-titanic_test
df2<-data.frame(Survived=rep("None",nrow(df2)),df2[,])#add survived column
df<-rbind(df1, df2)

sum(is.na(df$Embarked))
df$Embarked[is.na(df$Embarked)] <- 'S'
sum(is.na(df$Embarked))

sum(is.na(df$Age))

```

263 missing ages in 1309 obs

```{r}
sum(is.na(df$Age))
#263 na values
age_subset <-  df[complete.cases(df$Age), ] 
#all the ages without na values

#lets get rid of outliers

boxplot(age_subset$Age, horizontal = TRUE, axes = TRUE, staplewex = 1)
text(x = boxplot.stats(age_subset$Age)$stats, labels = boxplot.stats(age_subset$Age)$stats, y = 1.25)

clean_df<-age_subset[age_subset$Age <=66, ] 
#now i took out the outliers and the na values

```


```{r}
library(leaps)
bestsub.model <- regsubsets(Age~Pclass + 
                              Sex + Age + 
                              Fare+ SibSp + Parch+ Embarked, 
                            data = clean_df, nvmax =8)

summary(bestsub.model)

  cbind( 
    Cp     = summary(bestsub.model)$cp,
    r2     = summary(bestsub.model)$rsq,
    Adj_r2 = summary(bestsub.model)$adjr2,
    BIC    =summary(bestsub.model)$bic
)
```

i am not including survived bc it is not a valid predictor for age

CP - 5,6
R^2 - 5-7
R^2ADJ - 6
BIC - 3

CP - Mallows's Cp is used to assess the fit of a regression model that has been estimated using ordinary least squares. It is applied in the context of model selection, where a number of predictor variables are available for predicting some outcome, and the goal is to find the best model involving a subset of these predictors. A small value of Cp means that the model is relatively precise

R^2 - how much variance can be explained by the model

ADJUSTED R^2 - Adjusted R-squared is a modified version of R-squared that has been adjusted for the number of predictors in the model. The adjusted R-squared increases when the new term improves the model more than would be expected by chance. It decreases when a predictor improves the model by less than expected.

BIC - Bayesian information criterion aka Schwarz information criterion (also SIC, SBC, SBIC) is a criterion for model selection among a finite set of models; models with lower BIC are generally preferred

6 predictos: Pclass Sexmale SibSp Parch EmbarkedQ EmbarkedS
translates to these 5: Pclass Sex SibSp Parch Embarked

```{r}
reg.model = lm(Age~Pclass +Sex +SibSp+ Parch +Embarked, data = clean_df) 
summary(reg.model) 

```

```{r}

reg.model2 = lm(Age~Pclass +Sex +SibSp+Embarked, data = clean_df) 
summary(reg.model2) 
```
parch removed due to significance and complexity

now we have a model

```{r}
sum(is.na(df$Age))
summary(df$Age)

#take na values with columns i want
na_ages<-df[is.na(df$Age), c("Pclass","Sex","SibSp","Embarked", "Age")]

predicted_ages<-predict(reg.model2, newdata = na_ages)

summary(df$Age)




```


```{r}
df[is.na(df$Age),"Age"]<-predicted_ages
summary(df$Age)
sum(is.na(df$Age))


write.csv(df, "/Users/cordeliaradin/Desktop/datasets/age_complete_titanic.csv", row.names=FALSE)
```



