
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(vtable)
library(aod)
library(ggplot2)
library(leaps) 
library(tidyr)

#import dataset via console
df<-titanic_train

ncol(df)
colnames(df)
st(df)

```


```{r}
df %>%
  group_by(Survived) %>%
  summarize(n())

df %>%
  group_by(Survived, Pclass) %>%
  summarize(n())
```
regarding the binary question of whether or not someone would survive this event, and which factors were the greatest predictors - a logistic regression is the definitive choice
```{r}
sapply(df, sd)#stddev
#na
sum(is.na(df$Age))

df_age_edited<-df %>% drop_na(Age)
sapply(df_age_edited, sd)#stddev
sd(df_age_edited$Age)
```
```{r}
xtabs(~Survived+Sex+Pclass, data = df)

#use all of the potential vaiables to select the model with the best subset
bestsub.model <- regsubsets(Survived ~ Pclass + 
                              Sex + Age + 
                              Fare+ SibSp + Parch, 
                            data = df_age_edited, nvmax =6)

summary(bestsub.model)
```
that wasn't especially helpful in eliminating choices so lets look at other criteria

(also i tried treating Pclass like a catergorical variable as.factor at this stage and did not impact the model)

CP - Mallows's Cp is used to assess the fit of a regression model that has been estimated using ordinary least squares. It is applied in the context of model selection, where a number of predictor variables are available for predicting some outcome, and the goal is to find the best model involving a subset of these predictors. A small value of Cp means that the model is relatively precise

R^2 - how much variance can be explained by the model

ADJUSTED R^2 - Adjusted R-squared is a modified version of R-squared that has been adjusted for the number of predictors in the model. The adjusted R-squared increases when the new term improves the model more than would be expected by chance. It decreases when a predictor improves the model by less than expected.

BIC - Bayesian information criterion aka Schwarz information criterion (also SIC, SBC, SBIC) is a criterion for model selection among a finite set of models; models with lower BIC are generally preferred
```{r}


  cbind( 
    Cp     = summary(bestsub.model)$cp,
    r2     = summary(bestsub.model)$rsq,
    Adj_r2 = summary(bestsub.model)$adjr2,
    BIC    =summary(bestsub.model)$bic
)

```
BEST:
CP 4
R^2 4-6
AR^2 4
BIC 4

Model with 4 predictors is optimum
Pclass, Age, Sex, SibSp(#sibs spouse on ship)



```{r}
logistic_reg.model <- glm(Survived ~ Pclass + Sex + Age + SibSp, data = df_age_edited, family = "binomial")

summary(logistic_reg.model)
```
The indicator variables for passenger class are interpreted by class: if a passenger were in 2nd class accomadations vs being in 1rst class, changes the log odds of survival by -1.317398

If a passenger identifies as a male, this decreases the log odds of survival by -2.623483

For every additional year in Age, the log odds of survival (vs death) decreases by -0.044385

And for every additional sibling or spouse on the ship, decrease log odds of survival by -0.376119

```{r}
#confidence intervals for each of the variables
confint(logistic_reg.model)

wald.test(b = coef(logistic_reg.model), Sigma = vcov(logistic_reg.model), Terms = 2:5)
```
wald test with all variables results in a pvalue of 0, therefore all are significant

```{r}
exp(coef(logistic_reg.model))

exp(cbind(OR = coef(logistic_reg.model), confint(logistic_reg.model)))
```
odds ratio
```{r}
#lets treat passenger class as a categorical variable now
df_age_edited$Pclass <- factor(df_age_edited$Pclass)

logistic_reg.model2 <- glm(Survived ~ Pclass + Sex + Age + SibSp, data = df_age_edited, family = "binomial")

exp(coef(logistic_reg.model2))
```

```{r}
summary(logistic_reg.model)
```

```{r}
summary(logistic_reg.model2)#model with Pclass as catgorical
```
this really teases out the differences in passenger class

now probabilities

```{r}
df_age_edited2<-df_age_edited

newdata1 <- with(df_age_edited2, data.frame(Sex=Sex, Age = mean(Age), SibSp = mean(SibSp), Pclass = factor(1:3)))

newdata1

newdata1$rankP <- predict(logistic_reg.model2, newdata = newdata1, type = "response")
newdata1
```

holding age, sibsp as mean, here are the probabilities of survival for each class:
1rst class F: 0.94321328
1rst class M: 0.54545194 
2nd class F: 0.80148956
2nd class M: 0.22582306 
3rd class F: 0.53926206
3rd class M: 0.07796618 

```{r}

ggplot(newdata1, aes(x=Sex, y=rankP)) + 
  geom_point(alpha=.5) 

```

there is no R2 value for logistic regression. Instead, we can compute a metric known as McFadden’s R2, which ranges from 0 to just under 1. Values close to 0 indicate that the model has no predictive power. In practice, values over 0.40 indicate that a model fits the data very well.

```{r}
library(pscl)
pscl::pR2(logistic_reg.model2)["McFadden"]

```

```{r}
library(caret)

test_df<-titanic_test
test_df$Pclass <- factor(test_df$Pclass)
test_df<-test_df %>% drop_na(Age)


predicted <- predict(logistic_reg.model2, test_df, type="response")


predicted <- predict(logistic_reg.model2, df_age_edited2, type="response")



c1<-df_age_edited2$Survived
c2<-predicted
c2[c2 < .5] = 0
c2[c2 >= .5] = 1
c1<-as.factor(c1)#get thrown levels error if dont make as factor even though same type
c2<-as.factor(c2)

#Creating confusion matrix
example <- confusionMatrix(c1,c2)

#Display results 
example

```

here on the diagonals we can see that we had 361 true negative and 212 true positive.
and then we have our types 1&2
False Positive: Type 1 Error (predicted survive and they died)
78
False Negative: Type 2 Error (predicted dead and they survived)
63



